Generating gaits for physical quadruped robots: evolved neural networks vs. local parameterized search

### 2011-04-12 paper_ecal/experimentalSetup.tex
More edits

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world, instead of
using a simulator.  In this section we give details of the hardware
setup we used: \secref{platformDetails} discusses the robotic
platform, and \secref{fitnessEvaluation} describes the tracking system
we used to measure each gait's fitness.

>>>>

An important aspect of this study was to evaluate gait generation and
learning algorithms in the noisy, unforgiving real world instead of
using a simulator.  In this section we give details of the hardware
setup we used: \secref{platformDetails} discusses the robotic
platform, and \secref{fitnessEvaluation} describes the tracking system
we used to measure each gait's fitness.

### 2011-04-12 paper_ecal/hyperNeatMethod.tex
More edits

The ANN configuration follows previous studies that evolved quadruped
gaits with HyperNEAT in simulation~\cite{clune2011performance,
  clune2009evolving}, but was adapted to accommodate the physical robot used in this paper. Specifically, the ANNs consists of three $3 \times 4$
Cartesian grids of nodes forming input, hidden, and output
layers (\figref{SpiderANN.jpg}). Adjacent layers were allowed to be completely connected, meaning that there
could be $(3 \times 4)^2= 288$ links in each ANN (although evolution can set weights to 0, functionally eliminating the connection). The inputs to the
substrate were the current angles of each of the 9 joints of the robot and a sine and cosine wave (to facilitate the
production of periodic behaviors). The sine and cosine waves had a period of about half a second.

>>>>

The ANN configuration follows previous studies that evolved quadruped
gaits with HyperNEAT in simulation~\cite{clune2011performance,
  clune2009evolving}, but was adapted to accommodate the physical robot used in this paper. Specifically, the ANNs consists of three $3 \times 4$
Cartesian grids of nodes forming input, hidden, and output
layers (\figref{SpiderANN.jpg}). Adjacent layers were allowed to be completely connected, meaning that there
could be $(3 \times 4)^2= 288$ links in each ANN (although evolution can set weights to 0, functionally eliminating the connection). The inputs to the
substrate were the current \emph{commanded} angles of each of the 9 joints of the robot and a sine and cosine wave (to facilitate the
production of periodic behaviors).  The sine and cosine waves had a period of about half a second.

While measured angles could have been used as input instead of
commanded angles, this may have given HyperNEAT gaits an unfair
advantage by making them closed-loop. Thus, we restricted inputs to
the commanded angles only, which allowed us to produce completely
deterministic gaits using HyperNEAT.

### 2011-04-12 paper_ecal/introduction.tex
More edits

In this paper we compare the performance of two different methods of
learning gaits: parameterized gaits optimized with six different
learning methods, and gaits generated by evolving neural networks with
the HyperNEAT generative encoding~\cite{stanley2009hypercube}. While
some of these methods, such as HyperNEAT, have been tested in
simulation~\cite{clune2009evolving, clune2011performance}, we
investigate how they perform when evolving on a physical robot
(\figref{robot_whitebg.jpg}).

>>>>

In this paper we compare the performance of two different methods of
learning gaits: parameterized gaits optimized with six different
learning methods, and gaits generated by evolving neural networks with
the HyperNEAT generative encoding~\cite{stanley2009hypercube}. While
some of these methods, such as HyperNEAT, have been tested in
simulation~\cite{clune2009evolving, clune2011performance}, we
investigate how they perform when evolving on a physical robot
(\figref{robot_whitebg.jpg}) on flat terrain.

### 2011-04-12 paper_ecal/platformDetails.tex
More edits

\acmFig{topdown.png}{.7}{A figure of the robot from a top-down perspective,
with servos labeled}

The quadruped robot used in this study was assembled from parts
printed on the Objet Connex 500 3-D Printing System. It weights 1.88
kg with the on-board computer and measures approximately 38
centimeters from leg to opposite leg in the crouch position depicted
in \figref{robot_close.jpg}. The robot is actuated by 9 AX-12+
Dynamixel servos: one inner joint and one outer joint servo in each of
the four legs, and one servo at the center ``hip'' joint.  This final
unique servo allows the two halves of the robot to rotate with respect
to each other. \figref{topdown.png} shows the positions and numerical
designations of all nine servos.  Each servo could be commanded to a
position in the range [0,~1023], corresponding to a physical range
[-120$^{\circ}$,~+120$^{\circ}$].

All of the computation for gait learning, fitness evaluation, and
robot control was performed on the compact on-board CompuLab Fit-PC2,
running Ubuntu Linux.  All gait generation, learning, and fitness
evaluation code, except HyperNEAT, was written in Python and is
available on our website \cite{quadratotWeb}.  HyperNEAT is written in
C++.  We controlled the servos with the \code{pydynamixel} library
\cite{pydynamixel}.  The robot connected to a wireless network on
boot, which enabled us to control it via SSH.

>>>>

\acmFig{topdown2.jpg}{1}{(a) Top-down perspective of the robot with
  the nine joints and associated servos labeled. (b) The robot in a
  flat pose with the hip joint centered. (c,d,e) Various views of a
  pose in which the hip joint is rotated.}

The quadruped robot used in this study was assembled from parts
printed on the Objet Connex 500 3-D Printing System. It weighs 1.88 kg
with the on-board computer and measures approximately 38 centimeters
from leg to opposite leg in the crouch position depicted in
\figref{robot_whitebg.jpg}. The robot is actuated by 9 AX-12+
Dynamixel servos: one inner joint and one outer joint servo in each of
the four legs, and one servo at the center ``hip'' joint.  This final
unique servo allows the two halves of the robot to rotate with respect
to each other. \figref{topdown2.jpg} shows this unique motion, as well
as the positions and numerical designations of all nine servos.  Each
servo could be commanded to a position in the range [0,~1023],
corresponding to a physical range [-120$^{\circ}$,~+120$^{\circ}$].

All of the computation for gait learning, fitness evaluation, and
robot control was performed on the compact, on-board CompuLab Fit-PC2,
running Ubuntu Linux 10.10. 
The most computationally expensive portion of code was HyperNEAT, but even
this took less than one second to run on the on-board computer. Thus,
we chose not to offload any computation.
All gait generation, learning, and fitness
evaluation code, except HyperNEAT, was written in Python and is
available on our website \cite{quadratotWeb}.  HyperNEAT is written in
C++.  We controlled the servos with the \code{pydynamixel} library
\cite{pydynamixel}, sending commanded positions at 40Hz.  The robot
connected to a wireless network on boot, which enabled us to control
it via SSH.
